BOLD Parameter Estimation Using Particle Filters

### 2010-07-19 doc/thesis/results.tex
mostly done, just a bit of cleanup
need to go back and add some stuff to the slice sim

\section{Single Time Series}
Given the state-space equations for the BOLD signal, simulating a single time 
series is relatively straight forward. After a "true" signal is generated,
identically and independently distributed (I.I.D.) Gaussian noise and a Wiener
process with Gaussian I.I.D. steps are added to the true signal. Finally a 
"carrier level" is added, since BOLD is typically
measured as a \% difference from the baseline. The particle filter
algorithm will immediately remove this by calculating the \% difference, 
but adding a carrier level means that the exact same algorithm which 
will is used for simulated data may be used for the real data.

@@

%LOW NOISE SECTION, with a signal
\subsection{Signal with Low Noise}
%tests had noise of $ \{\sigma_y = .01, \sigma_x = .005\} $, where $\sigma_y$ is the
%measurement noise, and $\sigma_x$ is the wiener step size. Both signals used the
%parameters .
%The particle filter used the parameters defined in \autoref{tab:Prior} (\autoref{sec:PriorDistrib}),
%thus the particle filter was not centered over the correct values. 
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/realization_lownoise}
\caption{Test Signals with low noise compared to the clean signal.}
\label{fig:LowNoiseRealization}
\end{figure}

@@

\section{Multi-voxel Simulation}
To test the usefullness of the particle filter for Second I used a modified version of the FSL tool 
POSSUM to generate an entire FMRI image from a parameter map. The parameter map was generated
by taking an existing activation map and assigning discrete parameter sets to each region.
The result was a four dimensional (length x width
x height x parameter) image with spatially varying parameters. Possum was then modified
to take a parameter map and generate activation levels depending on the parameters at that
point. The patch for POSSUM will be made available. As an unfortunate side effect of 
not using Possums' original activation scale, I manually added 750 to the total level of
simulated Possum images. This is because the BOLD \% difference levels were in the range
of 50 - 100\% from the base, about 5 times as large as they should have been. Ultimately
this should not have an effect on the parameters (other than perhaps $\epsilon$ and $V_0$). 
For each time-series in the simulated FMRI image, the final \emph{static} parameters were saved
into a parameter map. This parameter map may then be compared to the map used to generate the 
simulated data; additionally a new simulation using the calculated parameters may also be 
generated to test the difference in activation levels between the real parameters and the
estimated ones. Since it is clear that the parameters are not fully orthogonal 
(\cite{Deneux2006}), its possible that two sets of parameters are functionally equivalent,
but have different parameters. This way, an absolute 
quantitative difference between the two parameter sets may be found.

>>>>

\section{Single Time Series}
\label{sec:Single Voxel Simulation}
Given the state-space equations for the BOLD signal, simulating a single time 
series is relatively straight forward. After a "true" signal is generated,
identically and independently distributed (I.I.D.) Gaussian noise and a Wiener
process with Gaussian I.I.D. steps are added to the true signal. Finally a 
"carrier level" is added, since BOLD is typically
measured as a \% difference from the baseline. The particle filter
algorithm will immediately remove this by calculating the \% difference, 
but adding a carrier level means that the exact same algorithm which 
will is used for simulated data may be used for the real data.

@@

%LOW NOISE SECTION, with a signal
\subsection{Signal with Low Noise}
\label{sec:SimLowNoise}
%tests had noise of $ \{\sigma_y = .01, \sigma_x = .005\} $, where $\sigma_y$ is the
%measurement noise, and $\sigma_x$ is the wiener step size. Both signals used the
%parameters .
%The particle filter used the parameters defined in \autoref{tab:Prior} (\autoref{sec:PriorDistrib}),
%thus the particle filter was not centered over the correct values. 
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/realization_lownoise}
\caption{Test Signals with low noise compared to the clean signal.}
\label{fig:LowNoiseRealization}
\end{figure}

@@

\section{Multi-voxel Simulation}
To test the usefullness of the particle filter for Second I used a modified version of the FSL tool 
POSSUM to generate an entire FMRI image from a parameter map. The parameter map was generated
by taking an existing activation map and assigning discrete parameter sets to each region.
The result was a four dimensional (length x width
x height x parameter) image with spatially varying parameters. Possum was then modified
to take a parameter map and generate activation levels depending on the parameters at that
point. The patch for POSSUM will be made available. As an unfortunate side effect of 
not using Possums' original activation scale, I manually added 750 to the total level of
simulated Possum images. This is because the BOLD \% difference levels were in the range
of 50 - 100\% from the base, about 5 times as large as they should have been. Ultimately
this should not have an effect on the parameters (other than perhaps $\epsilon$ and $V_0$). 
Also worth noting is that the noise level was set to an SNR of 20, due to changes in the 
program the true signal-to-noise ratio was much lower, as soon in the fit in \autoref{fig:simslicefit1}.

\begin{figure}
%includegraphics[]{images/simslicefit1}
\caption{A single time-series fit in the simulated slice.}
\labe{fig:simslicefit1}
\end{figure}

For each time-series in the simulated FMRI image, the final \emph{static} parameters were saved
into a parameter map. This parameter map may then be compared to the map used to generate the 
simulated data; additionally a new simulation using the calculated parameters may also be 
generated to test the difference in BOLD levels between the real parameters and the
estimated ones. Since it is clear that the parameters are not fully orthogonal 
(\cite{Deneux2006}), its possible that two sets of parameters are functionally equivalent,
yet different. This way, a quantitative difference between the two parameter sets 
may be found.

\begin{figure}
\includegraphics[]{images/simslice_hm}
\caption{Normalized Square Root of the Mean Square Residual. Darker indicates lower error.}
\label{fig:simslice_hm}
\end{figure}

\begin{figure}
\includegraphics[]{images/simslice_hm85}
\caption{Normalized Square Root of the Mean Square Residual, with upper threshold at .85. 
Darker indicates lower error.}
\label{fig:simslice_hm85}
\end{figure}

\begin{figure}
\includegraphics[]{images/parameters_sec1}
\caption{Histogram of estimated parameters in section with true parameters of $\{ \}$
Limited to areas with residual less than $.85$.}
\label{fig:parameters_sec1}
\end{figure}

\begin{figure}
\includegraphics[]{images/parameters_sec2}
\caption{Histogram of estimated parameters in section with true parameters of $\{ \}$
Limited to areas with residual less than $.85$.}
\label{fig:parameters_sec2}
\end{figure}

\begin{figure}
\includegraphics[]{images/parameters_sec3}
\caption{Histogram of estimated parameters in section with true parameters of $\{ \}$. 
Limited to areas with residual less than $.85$.}
\label{fig:parameters_sec3}
\end{figure}
